### 2023-01-16 latex/includes/pressure_vessel_problem.tex
Analysis and latex formatting quite done

\begin{table}
\centering
\caption{Statistical Information about function values for Pressure Vessel Design}
\label{function_values:pressure_vessel_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      Min F &        Mean F &      Median F &          Max F &      StdDev F \\
\midrule
       EP & 0.94974179 &    0.99959274 &    0.95597819 &     1.41652156 &    0.10256309 \\
       ES & 0.94985828 &    3.07906532 &    1.90278615 &    11.41620650 &    2.47339648 \\
       GA & 4.61928311 &  383.23040981 &  262.27555979 &  1747.41758530 &  362.98174772 \\
    BeesA & 1.05944833 & 5212.47847291 & 2094.42824655 & 30412.73885965 & 7297.64717603 \\
      FFA & 0.95407161 &    1.05749808 &    1.05432826 &     1.16958618 &    0.04485239 \\
      PSO & 0.94936617 &    1.00652465 &    0.94974267 &     3.01988559 &    0.25222125 \\
\bottomrule
\end{tabular}
\end{table}