### 2023-01-16 latex/includes/spring_problem.tex

\begin{table}
\centering
\caption{Statistical Information about function values for Spring Tension Design}
\label{function_values:spring_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      Min F &     Mean F &   Median F &      Max F &   StdDev F \\
\midrule
       EP & 0.00607563 & 0.00607590 & 0.00607570 & 0.00607834 & 0.00000044 \\
       ES & 0.00607564 & 0.00609326 & 0.00608915 & 0.00616308 & 0.00001810 \\
       GA & 0.00610989 & 0.00670843 & 0.00663331 & 0.00805670 & 0.00045944 \\
    BeesA & 0.00607781 & 0.00632310 & 0.00625074 & 0.00724816 & 0.00025469 \\
      FFA & 0.00607563 & 0.00607593 & 0.00607576 & 0.00607847 & 0.00000047 \\
      PSO & 0.00607563 & 0.00607563 & 0.00607563 & 0.00607563 & 0.00000000 \\
\bottomrule
\end{tabular}
\end{table}