### 2023-01-16 latex/includes/pressure_vessel_problem_boxplot.tex

\begin{figure}[H]
\centering
\caption{Boxplot for Pressure Vessel Design}
\label{fig:pressure_vessel_design_boxplot}
\includegraphics[scale=0.5]{images/pressure_vessel_problem_boxplot.png}
\end{figure}