### 2023-01-19 latex/includes/pressure_vessel_problem.tex

\begin{table}
\centering
\caption{Statistical Information about function values for Pressure Vessel Design}
\label{function_values:pressure_vessel_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &         Min F &        Mean F &      Median F &         Max F &   StdDev F \\
\midrule
       EP & 5534.57345496 & 5534.63426098 & 5534.61416291 & 5534.92993607 & 0.05804548 \\
       ES & 5534.57917966 & 5535.28165367 & 5535.06149254 & 5538.03003895 & 0.65098467 \\
       GA & 5534.83662424 & 5538.36788889 & 5536.92932748 & 5550.91639342 & 3.57079951 \\
    BeesA & 5534.57516657 & 5537.97495231 & 5537.04824684 & 5548.55587234 & 3.28239740 \\
      FFA & 5534.57401617 & 5534.58849395 & 5534.58435493 & 5534.63296149 & 0.01341777 \\
      PSO & 5534.57345480 & 5534.57381593 & 5534.57355270 & 5534.57625290 & 0.00053848 \\
\bottomrule
\end{tabular}
\end{table}