### 2023-01-19 latex/includes/pressure_vessel_problem_original.tex

\begin{table}
\centering
\caption{Statistical Information about function values for Pressure Vessel Design (Original)}
\label{function_values:pressure_vessel_problem_original}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &         Min F &         Mean F &       Median F &          Max F &       StdDev F \\
\midrule
       EP &    0.08150806 &     0.15642850 &     0.09093798 &     1.32637552 &     0.16807714 \\
       ES &    0.08161190 &   718.92910990 &     1.00278859 &  9443.54269260 &  2060.07774212 \\
       GA &   17.06920868 &   525.00734162 &   361.53879602 &  3112.03503082 &   454.52386703 \\
    BeesA & 2393.70106213 & 17564.33024022 & 15751.91467117 & 53850.38271384 & 10514.60780411 \\
      FFA &    0.08306113 &     0.13296127 &     0.13106271 &     0.20084847 &     0.02396876 \\
      PSO &    0.08149204 &     0.24191732 &     0.08159173 &    10.00611523 &     1.00072349 \\
\bottomrule
\end{tabular}
\end{table}