### 2023-01-19 latex/includes/pressure_vessel_problem_convergence.tex
Formatting Done.

\begin{figure}[H]
\centering
\caption{Convergence lines for Pressure Vessel Design}
\label{fig:pressure_vessel_problem_convergence}
\includegraphics[width=0.4 \textwidth]{images/pressure_vessel_problem_convergence.png}
\end{figure}