### 2023-01-19 latex/includes/pressure_vessel_problem_original_convergence.tex

\begin{figure}[H]
\centering
\caption{Convergence lines for Pressure Vessel Design (Original)}
\label{fig:pressure_vessel_problem_original_convergence}
\includegraphics[width= 0.4 \textwidth]{images/pressure_vessel_problem_original_convergence.png}
\end{figure}