### 2023-01-20 latex/includes/pressure_vessel_problem_convergence.tex

Figure~\ref{fig:pressure_vessel_problem_convergence} shows the
function maximum, minimum and mean function from algorithm's evolution:

\begin{figure}[H]
\centering
\caption{Convergence lines for Pressure Vessel Design}
\label{fig:pressure_vessel_problem_convergence}
\includegraphics[width=0.4 \textwidth]{images/pressure_vessel_problem_convergence.png}
\end{figure}

EP, ES and PSO quickly converges all the solution once it find a viable minimum, whe GA, BeesA and FFA
maintain exloration and explotation for all the epochs. From Table~\ref{function_values:pressure_vessel_problem}
, PSO has the best solution stability for our proud-lame-homemade custom bounding problem.