### 2023-01-20 latex/includes/pressure_vessel_problem_original_convergence.tex

Figure~\ref{fig:pressure_vessel_problem_original_convergence} shows the
function maximum, minimum and mean function from algorithm's evolution:

\begin{figure}[H]
\centering
\caption{Convergence lines for Pressure Vessel Design (Original)}
\label{fig:pressure_vessel_problem_original_convergence}
\includegraphics[width=0.4 \textwidth]{images/pressure_vessel_problem_original_convergence.png}
\end{figure}

All algorithms tested have slow evolution for the original variable boundings, as it is shown in the convergence evolution.