### 2023-01-20 latex/includes/spring_problem.tex

Another view of the solutions are presented in Table~\ref{function_values:pressure_vessel_problem}, where
are summarized the statistics of all 100 possible solutions for different starting points (populations).

\begin{table}[H]
\centering
\caption{Statistical Information about function values for Spring Tension Design}
\label{function_values:spring_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      Min F &     Mean F &   Median F &      Max F &   StdDev F \\
\midrule
       EP & 0.01250534 & 0.01255960 & 0.01253156 & 0.01453933 & 0.00020190 \\
       ES & 0.01252956 & 0.01281857 & 0.01274283 & 0.01346214 & 0.00024181 \\
       GA & 0.01252305 & 0.01628575 & 0.01616768 & 0.02561808 & 0.00273276 \\
    BeesA & 0.01250200 & 0.01268010 & 0.01261691 & 0.01406679 & 0.00021629 \\
      FFA & 0.01249803 & 0.01254780 & 0.01253499 & 0.01268116 & 0.00003152 \\
      PSO & 0.01252626 & 0.01390866 & 0.01318774 & 0.03036537 & 0.00342240 \\
\bottomrule
\end{tabular}}
\end{table}