### 2023-01-21 release/code/latex/includes/pressure_vessel_problem.tex

Another view of the solutions are presented in Table~\ref{function_values:pressure_vessel_problem}, where
are summarized the statistics of all 100 possible solutions for different starting points.

\begin{table}[H]
\centering
\caption{Statistical Information about function values for Pressure Vessel Design}
\resizebox{\columnwidth}{!}{%
\label{function_values:pressure_vessel_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &         Min F &        Mean F &      Median F &         Max F &   StdDev F \\
\midrule
       EP & 5534.57345496 & 5534.63426098 & 5534.61416291 & 5534.92993607 & 0.05804548 \\
       ES & 5534.57917966 & 5535.28165367 & 5535.06149254 & 5538.03003895 & 0.65098467 \\
       GA & 5534.83662424 & 5538.36788889 & 5536.92932748 & 5550.91639342 & 3.57079951 \\
    BeesA & 5534.57516657 & 5537.97495231 & 5537.04824684 & 5548.55587234 & 3.28239740 \\
      FFA & 5534.57401617 & 5534.58849395 & 5534.58435493 & 5534.63296149 & 0.01341777 \\
      PSO & 5534.57345480 & 5534.57381593 & 5534.57355270 & 5534.57625290 & 0.00053848 \\
\bottomrule
\end{tabular}}
\end{table}

It is hard to see any statistical  difference looking at any value in the table except standard deviation.
GA and BeesA seems to have the worst solutions, and EP, FFA and PSO seems to have a well defined behavior
since they have the lowest standard deviations, that is, all the solutions don't spread too much from each
other.