### 2023-01-21 release/code/latex/includes/pressure_vessel_problem_best_fits.tex

In this section are presented the results for a tweaked variable boundings of Pressure Vessel Design,
as exposed in Section~\ref{subsubsec:methodology-pressure-vessel-design}.

The best results for all algorithms are presented in Table~\ref{best_fits:pressure_vessel_problem}:

\begin{table}[H]
\centering
\caption{Best Fits for Pressure Vessel Design}
\label{best_fits:pressure_vessel_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      $x_1$ &      $x_2$ &       $x_3$ &        $x_4$ &         $f_x$ \\
\midrule
       EP & 0.75000000 & 0.35000000 & 40.68318345 & 195.00000000 & 5534.57345496 \\
       ES & 0.75000000 & 0.35000000 & 40.68322151 & 195.00000000 & 5534.57917966 \\
       GA & 0.75000816 & 0.35003661 & 40.68379662 & 195.00019653 & 5534.83662424 \\
    BeesA & 0.75000000 & 0.35000000 & 40.68319525 & 195.00000000 & 5534.57516657 \\
      FFA & 0.75000000 & 0.35000000 & 40.68318772 & 195.00000000 & 5534.57401617 \\
      PSO & 0.75000000 & 0.35000000 & 40.68318348 & 195.00000000 & 5534.57345480 \\
\bottomrule
\end{tabular}}
\end{table}

It is noticeable in this table that the best results and the best minimizers are quite the same,
so it cannot be said that all algorithms will solve the problem from this standpoint.