### 2023-01-21 release/code/latex/includes/pressure_vessel_problem_original_best_fits.tex

In this section are presented the results for the original variable boundings of Pressure Vessel Design,
as exposed in Section~\ref{subsubsec:methodology-pressure-vessel-design}.

The best results for all algorithms are presented in Table~\ref{best_fits:pressure_vessel_problem_original}:

\begin{table}[H]
\centering
\caption{Best Fits for Pressure Vessel Design (Original)}
\label{best_fits:pressure_vessel_problem_original}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      $x_1$ &      $x_2$ &        $x_3$ &        $x_4$ &         $f_x$ \\
\midrule
       EP & 0.00000000 & 0.00000000 &  40.32041055 & 200.00000000 &    0.08150806 \\
       ES & 0.00000000 & 0.00000000 &  40.32554414 & 200.00000000 &    0.08161190 \\
       GA & 0.00337818 & 0.00106498 &  46.44471109 & 129.50295637 &   17.06920868 \\
    BeesA & 0.07617198 & 0.07091908 & 104.75191725 &  44.23570263 & 2393.70106213 \\
      FFA & 0.00000000 & 0.00000000 &  40.39658670 & 200.00000000 &    0.08306113 \\
      PSO & 0.00000000 & 0.00000000 &  40.31961791 & 200.00000000 &    0.08149204 \\
\bottomrule
\end{tabular}}
\end{table}

While EP, ES, FFA and PSO seems to give similar results, it is noticeable that the cost function
gave some discrepant results for GA and BeesA.