### 2023-01-21 release/code/latex/includes/spring_problem_best_fits.tex

In this section are presented the results of Spring Tension Design optimization,
as exposed in Section~\ref{subsubsec:methodology-spring-tension-design}.

The best results for all algorithms are presented in Table~\ref{best_fits:spring_problem}:

\begin{table}[H]
\centering
\caption{Best Fits for Spring Tension Design}
\label{best_fits:spring_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrr}
\toprule
Algorithm &      $x_1$ &      $x_2$ &       $x_3$ &      $f_x$ \\
\midrule
       EP & 0.05078716 & 0.34120356 & 11.93931414 & 0.01250534 \\
       ES & 0.05000000 & 0.32292194 & 13.26929635 & 0.01252956 \\
       GA & 0.05260626 & 0.38419911 &  9.63922108 & 0.01252305 \\
    BeesA & 0.05141008 & 0.35564113 & 11.06109433 & 0.01250200 \\
      FFA & 0.05148403 & 0.35770175 & 10.95787946 & 0.01249803 \\
      PSO & 0.05000000 & 0.32241990 & 13.31368880 & 0.01252626 \\
\bottomrule
\end{tabular}}
\end{table}

Although the best fit points have noticeable differences, the cost function $f_x$ have similar results, that is, the
function $f_x$ seems to have weak minimal points.