### 2023-01-21 release/code/latex/includes/spring_problem_convergence.tex

Figure~\ref{fig:spring_problem_convergence} shows the
function maximum, minimum and mean function of algorithm's evolution:

\begin{figure}[H]
\centering
\caption{Convergence lines for Spring Tension Design}
\label{fig:spring_problem_convergence}
\includegraphics[width=0.4 \textwidth]{images/spring_problem_convergence.png}
\end{figure}

EP, ES and FFA have the best fitness.
All algorithms seems to converge the minimum line very quickly when
the algorithm find a minimal point.
Apart from EP and ES, the algorithms seems to maintain exploration and exploitation
until the time runs out.