### 2023-01-21 release/code/latex/includes/spring_problem_wilcoxon_test.tex

The Table~\ref{wilcoxon_test:spring_problem} gives the Wilcoxon test results
for \textit{Spring Tension Design} as exposed in Section~\ref{subsubsec:methodology-spring-tension-design}.

\begin{table}[H]
\centering
\caption{Significance Test Using Wilcoxon Test for Spring Tension Design}
\label{wilcoxon_test:spring_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lllllll}
\toprule
  --- & EP & ES & GA & BeesA & FFA & PSO \\
\midrule
   EP & 0.0 & 1.47e-16 & 4.01e-18 & 4.37e-12 & 0.01 & 2.33e-16 \\
   ES & 1.47e-16 & 0.0 &  1.90e-17 & 2.12e-06 & 3.24e-17 & 7.98e-12 \\
   GA & 4.01e-18 &  1.90e-17 & 0.0 & 8.02e-18 & 4.27e-18 & 1.29e-11 \\
BeesA & 4.36e-12 & 2.12e-06 & 8.02e-18 & 0.0 & 3.01e-11 & 6.82e-15 \\
  FFA & 0.01 & 3.23e-17 & 4.27e-18 & 3.01e-11 & 0.0 & 9.32e-18 \\
  PSO & 2.33e-16 & 7.98e-12 & 1.29e-11 & 6.82e-15 & 9.32e-18 & 0.0 \\
\bottomrule
\end{tabular}}
\end{table}

From Wilcoxon Test, the are no difference between all algorithms since p-value
is less than $0.05$.